---
title: "R Notebook"
output: html_notebook
---

# R Final Project

## Space Travel Data Set

Pulling in train.csv from Kaggle site.

```{r}
read_csv("train.csv")
sp_data <- as.data.frame(read_csv("train.csv"))

```

```{r}
class(sp_data)
```

### Data Types

-   Continuous

    -   Age
    -   Spending Categories
        -   RoomServices

        -   FoodCourt

        -   ShoppingMall

        -   Spa

        -   VRDeck

-   Discrete

    -   KEY- Passenger ID

    -   HomePlanet

    -   CryoSleep

    -   Cabin

    -   VIP

    -   Name

    -   Transported

## Query for Basic Information

### Group Sizes

```{r}
#What is the average group size?
group_info <- group_info %>%
             summarise(group_size = n())
avg_grp_size <- group_info %>% 
  summarise(avg = mean(group_size), count = n())
avg_grp_size
# avg group size is less than 1.5 out of 69217 groups
View(group_info_df
```

```{r}


ggplot(group_info_df, aes(x=group_size)) +
         geom_histogram(bins = 8, fill = "purple", col = "black") +
  labs(title="Groups by Count and Size", y= "Count of Groups", x="Group Size") +
my_theme
```

### VIP Count

```{r}
#How many of our passengers are VIPs
vip_info <- sp_data_w_age %>%
  group_by(VIP, HomePlanet)%>%
  summarise(count = n())
vip_info

ggplot(vip_info) +
  geom_col(x=HomePlanet, y = count, fill = VIP) + 
  scale_color_brewer("Dark2") +
  my_theme()

```

### Cryo Sleep Info

```{r}
#How many people opted for the Cryo sleep option (and how many of them are VIPs)
cryo_data <- sp_data_w_age %>%
  group_by(CryoSleep, traveler_age_group ) %>%
  summarise(count = n())
cryo_data #only 21 VIPs using cryosleep


```

```{r}

ggplot (sp_data) +
    geom_bar(mapping = aes(x= CryoSleep, fill= VIP, position = "dodge"), color = "black") +
  theme (aspect.ratio = 1) + scale_fill_brewer(palette = "Dark2") +
  labs(x = NULL, y = NULL) + my_theme + labs(title = "Number of CryoSleep Users by VIP Status", x= "Passengers", y= "CryoSleep")
```

### Transport Info

```{r}
#number of passengers transported by destination, home planet
transport_info <- sp_data %>%
  group_by(Destination, Transported) %>%
  filter(!is.na(Destination)) %>%
  summarise(num_passengers = n())
arrange(transport_info, desc(num_passengers))
```

```{r}
ggplot (sp_data) +
    geom_bar(mapping = aes(x= Transported, fill= Destination, position = "fill"), color = "black") +
  theme (aspect.ratio = 1) + coord_polar() + 
  scale_fill_manual(values=c("#009668","#eeff41", "#ffab40")) +
  labs(x = NULL, y = NULL) + my_theme + labs(title = "Number of Completed Trips by Planet")
```

### Destination Info

```{r}
dest_info <- sp_data %>%
  group_by(Destination) %>%
  filter(!is.na(Destination)) %>%
  summarise(num_passengers = n())
dest_info

```

```{r}
traveler_dest <- age_group %>% 
  group_by(Destination, traveler_age_group) %>% 
    filter(!is.na(Destination)) %>%
  summarise(
    num_passengers = n(),
    na.rm = TRUE
  )   
traveler_dest
```

## Query for demographics information

### Grouping Ages

```{r}
by_age <- sp_data %>% count(Age, na.rm=TRUE)
by_age_sorted <- arrange(by_age,desc(n))
by_age_sorted  <- by_age_sorted %>% filter(!is.na(Age))
by_age_sorted

#looks like making age groups will yield interesting info

#by_age_sorted %>% summarize(median_age = median(Age))

ggplot(sp_data_w_age) +
  geom_density(aes(Age)) +
  labs(title = "Passenger Age Distribution") +
   my_theme
```

```{r}
#find min & max ages
min_age <- min(sp_data$Age, na.rm = T)
min_age
max_age <- max(sp_data$Age, na.rm=T)
max_age
#age groups: 0-3 'baby', 4-16 'child',17-30 'young adult', 31-60 'adult','60+ senior'
```

```{r}
#add age group column to data set
age_group <- sp_data %>%
  group_by(Age) %>%
  mutate(traveler_age_group = case_when(
    Age >= 65 ~ 'senior',
    Age >= 31 ~ 'adult',
    Age >= 17 ~ 'young adult',
    Age >= 4  ~ 'child',
    Age >= 0  ~ 'baby',
    TRUE ~ NA_character_
  ))

sp_data_w_age <- as.data.frame(age_group) # making new results into DF, hopefully it will help with issues I've been having
age_group_sum <- age_group %>%
  group_by(traveler_age_group) %>%
  filter (!is.na(traveler_age_group)) %>%
  summarise(num_passengers = n())
age_group_sum
```

```{r}
age_pie = ggplot(data = age_group_sum) +
  geom_col(mapping = aes(x = traveler_age_group, y=num_passengers, fill = traveler_age_group,position = "dodge"), color= "black") +
  theme (aspect.ratio = 1) + 
  labs (title= "Traveler Age Groups", y="Number of Passengers", x = "Age Groups") +
  my_theme

age_pie 
```

### Home Planet

```{r}
traveler_home<- sp_data %>%
  group_by(HomePlanet) %>%
  filter(!is.na(HomePlanet)) %>%
  summarise(num_passengers = n(),
            na.rm = TRUE)
arrange(traveler_home, HomePlanet, desc(num_passengers))
#majority of travelers are from Earth
```

#### Home Planet by Age Group

```{r}
#taking a look at where age groups are coming from
traveler_home_age <- sp_data_w_age %>%
  group_by(traveler_age_group, HomePlanet) %>%
    filter(!is.na(HomePlanet)) %>%
  summarise(num_passengers = n(),
            na.rm = TRUE)
 View(arrange(traveler_home_age, desc(num_passengers), HomePlanet, traveler_age_group))
  
#most travelers are YOUNG adults from earth!
```

#### VIPs by Home Planet, Age

```{r}
VIP_age <- sp_data_w_age %>%
  group_by(VIP, traveler_age_group) %>%
  summarise(num_passengers = n())
VIP_age
```

```{r}
#backing up to see where our VIPs come from
VIP_home <- sp_data %>%
  filter( VIP == TRUE) %>%
  group_by(HomePlanet)%>%
  summarise(num_VIPs = n()) 
  

VIP_home <- as.data.table(VIP_home)# agreeing with the data below, all of our VIPs are from Europa & Mars
ggplot(total_spent) +
  geom_point(aes(x=HomePlanet, y=Age, color = VIP), alpha =.5, position = "jitter") +
  scale_size_area() +  my_theme
```

## Query for Financial Information

### Spending on Services

```{r}
#getting some totals and averages for spending
rs_ttl<-sum(sp_data$RoomService, na.rm = T)
fc_ttl<-sum(sp_data$FoodCourt, na.rm = T)
sm_ttl<-sum(sp_data$ShoppingMall, na.rm = T)
spa_ttl<-sum(sp_data$Spa, na.rm = T)
vr_ttl<-sum(sp_data$VRDeck, na.rm = T)

#number of customers
rs_cust <- sp_data %>% filter (RoomService > 0) %>% summarise(rs_cust = n()) %>% pull(rs_cust)
fc_cust <- sp_data %>% filter (FoodCourt > 0) %>% summarise(fc_cust = n()) %>% pull(fc_cust)
sm_cust <- sp_data %>% filter(ShoppingMall > 0) %>% summarise(sm_cust = n()) %>% pull(sm_cust)
spa_cust <- sp_data %>% filter (Spa > 0) %>% summarise(spa_cust = n()) %>% pull(spa_cust)
vr_cust <- sp_data %>% filter (VRDeck > 0) %>% summarise(vr_cust = n()) %>% pull(vr_cust)

service_name <- c("Room Service","Food Court","Shopping Mall","Spa","VR Deck")

service_spending<- c(rs_ttl, fc_ttl, sm_ttl, spa_ttl, vr_ttl)
avg_spending_per_traveler <- c(mean(rs_ttl/8693),mean(fc_ttl/8693),mean(sm_ttl/8693),mean(spa_ttl/8693), mean(vr_ttl/8693))

sdf <- data.frame(service_name, service_spending, avg_spending_pp)

#ttl_spending <- sum(sdf$service_spending)
#avg_spending <- mean(ttl_spending/8693)
#ttl_spending
#avg_spending

#arrange(sdf, service_spending)
sdf$customer_count <- c(rs_cust, fc_cust, sm_cust, spa_cust, vr_cust)
sdf$avg_per_customer <- c((rs_ttl/rs_cust),(fc_ttl/fc_cust),(sm_ttl/sm_cust),(spa_ttl/spa_cust),(vr_ttl/vr_cust))
sdf 
```

```{r}
ggplot(sdf, aes(x=service_name, y=avg_per_customer, fill = service_name)) +
  geom_col(color = "black") + my_theme +
  labs(title="Average Customer Spending per Vendor Type", x= "Vendor Type", y="Average Spent ($)", color = "Service Name") + guides(x = guide_axis(n.dodge = 2))
```

```{r}
#How much did individuals spend?
total_spent <- sp_data_w_age %>%
 mutate(ttl_spent=rowSums(across(c(RoomService:VRDeck)))) 

total_spent
#top spender was NOT a VIP
```

```{r}
ggplot(age_group_spending, aes(x= traveler_age_group, fill= factor(VIP))) +
  geom_dotplot(stackratio = .7, stackdir = "center", binpositions = "all") +
  scale_y_continuous(NULL, breaks = NULL)
```

#### VIPs

```{r}
#spending by VIP
VIP_spending <- total_spent %>%
  group_by(HomePlanet) %>%
  filter(VIP == TRUE) %>%
  summarise(spending = sum(ttl_spent, na.rm = TRUE), num_passengers = n(), avg_spending = spending/num_passengers)

VIP_spending #only VIPs from Europa & Mars spending $$
```

#### Age Groups

```{r}
#spending by age group
#having grouping issues

age_group_spending <- total_spent %>%
  group_by(Age, traveler_age_group,HomePlanet,VIP) %>%
  filter(!is.na(traveler_age_group))%>%
  summarize(spending = sum(ttl_spent, na.rm =TRUE), rs= sum(RoomService), fc = sum(FoodCourt), sm =sum(ShoppingMall), spa = sum(Spa), vr = sum(VRDeck), num_passengers = n(), avg_spending = spending/num_passengers)

arrange(age_group_spending, desc(avg_spending), traveler_age_group, spending, num_passengers) #on average seniors are spending the most
```

#### Home Planet

```{r}
spending_by_homeplanet <- total_spent %>%
  group_by(HomePlanet) %>%
  filter(!is.na(HomePlanet))%>%
  summarize(spending = sum(ttl_spent, na.rm =TRUE), num_passengers = n(), avg_spending = spending/num_passengers)

spending_by_homeplanet

ggplot(spending_by_homeplanet, aes(x = HomePlanet, y= spending)) +
  geom_point() +
geom_segment(aes(x = HomePlanet, xend = HomePlanet, y=0, yend=spending, color = HomePlanet), line =10) + my_theme
```

## Plotting Results

```{r}
#setting a theme
my_theme <- theme_minimal() + theme(text = element_text(face = "bold", family = "Courier New"))
```

### Home Planet Data

```{r}
h<-traveler_home_age  %>% 
  ggplot(mapping = aes(y=num_passengers, x=HomePlanet, 
                       fill=HomePlanet)) +
        labs(title = "Number of Passengers by Origin", y= "Passengers", x ="Home Planet") + 
        my_theme
h <- h + geom_boxplot() 
h
```

### Destination Planet

```{r}
d<-traveler_dest  %>% 
  ggplot(mapping = aes(y=num_passengers, x=Destination, fill=Destination)) +
  geom_boxplot()
d <- d + labs(title = "Number of Passengers by Destination", y = "Passengers", x = "Destination Planet") + 
  scale_fill_manual(values=c("#009668","#eeff41", "#ffab40")) + my_theme

d
```

```{r}
#
ggplot_add(d + h)
```

```{r}
#home planet, age
home_age<-traveler_home_age  %>% 
  ggplot(mapping = aes(y=num_passengers, x=traveler_age_group, fill= HomePlanet)) +
  geom_col() + scale_fill_brewer(palette = "Dark2") + labs(title = "Travelers by Age Group and Home Planet", y= "Passengers") + my_theme
  
home_age
```

```{r}
#another look at age/planet
age_group_drop <- traveler_home_age %>%
  filter(!is.na(traveler_age_group))

home_age_2 <- ggplot(sp_data_w_age, aes(traveler_age_group, fill = HomePlanet)) +
  geom_bar(position = "dodge", color="black") + scale_fill_brewer(palette = "Dark2") + labs(title = "Travelers by Age Group and Home Planet", y= "Passengers", x= "Traveler Age Group") + my_theme
           
#+ scale_fill_manual(values =c("#009668","#eeff41", "#ffab40")) + my_theme + facet_wrap(vars(traveler_age_group)
           
           
home_age_2
```

```{r}


dest_age_2 <- ggplot(sp_data_w_age, aes(traveler_age_group, fill = Destination)) +
  geom_bar(position = "dodge", color = "black") +  scale_fill_manual(values=c("#009668","#eeff41", "#ffab40")) + labs(title = "Travelers by Age Group and Destination", y= "Passengers", x= "Traveler Age Group") + my_theme
           
#+ scale_fill_manual(values =c("#009668","#eeff41", "#ffab40")) + my_theme + facet_wrap(vars(traveler_age_group)
           
           
dest_age_2
```

### Spending Data

```{r}
new_ags <- age_group_spending %>%
  filter(Age >= 17)

p <- ggplot(new_ags) +
  geom_hex (aes(x= num_passengers, y= spending, color= VIP, fill=HomePlanet)) +
  labs (x = "Passengers", y = "Total Expenditures ($)", title = "Space Traveler Expenses by Home Planet") +  scale_fill_manual(values=c("#009668","#eeff41", "#ffab40")) +
  my_theme
p
```

```{r}
pp <- ggplot(new_ags) +
  geom_hex (aes(x=num_passengers, y= spending, fill= traveler_age_group, color= VIP)) + 
  labs (x = "Number of Passengers", y = "Total Expenditures ($)", title = "Space Traveler Expenses by Age Group") + 
  scale_size_area() +
  my_theme
pp
```

```{r}
#Age & planet
ap <- ggplot(new_ags, aes(x=Age, y=spending)) +
               geom_point(aes(color = HomePlanet, size = num_passengers),alpha = .7)+
  scale_size_area(max_size = 15) + labs(title = "Spending Trends by Age, Home Planet")+my_theme
ap
```

```{r}
#plot spending data by service
ags <- new_ags [, c("traveler_age_group","rs","fc","sm","spa","vr")]


ags <- as.data.table(ags)
newer_ags <- melt(ags, id.vars = 'traveler_age_group', variable.name ='series')

agsp <- ggplot(newer_ags, aes(series, traveler_age_group)) +
  geom_jitter(aes(color=series)) +
  labs(title = "Service Customers by Age Group", x = "Traveler Age Group") +
  my_theme
agsp
```

```{r}
pie = ggplot(data = sp_data_w_age) +
  geom_bar(mapping = aes(x=VIP, fill = traveler_age_group),
           position = "fill"
           ) +
  theme (aspect.ratio = 1) +
  labs(x = NULL, y = NULL) + my_theme

pie + coord_polar() 

```

```{r}
pie = ggplot(data = sp_data_w_age) +
  geom_bar(mapping = aes(x=traveler_age_group, fill = Transported),
           position = "fill"
           ) + scale_fill_distiller(palette = "Greens")
  theme (aspect.ratio = 1) +
  labs(x = NULL, y = NULL) + my_theme

pie + coord_polar() 
```

```{r}

ggplot(total_spent, aes(x = ttl_spent)) +
  geom_histogram(binwidth = 10, fill = "#BDBDBD", col = "white", alpha = 0.7) +
  geom_histogram(data = , aes(x = dep_delay), binwidth = 10, fill = "#6A3D9A", col = "white", alpha = 0.7) +
  labs(
    title = "Distribution of Arrival and Departure Delays",
    x = "Delay (minutes)",
    y = "Frequency"
  ) +
  coord_cartesian(xlim = c(-100, 200)) +
  theme_dark()
```
